---
title             : "Effect of Gender and Socioeconomic Status on Math Score"
shorttitle        : "Research Article"

author: 
  - name          : "Jessica Canfield"
    corresponding : yes    # Define only one corresponding author
    address       : "1208 University of Oregon, Eugene, OR 97403"
    email         : "jcanfiel@uoregon.edu"
  - name          : "Woocheol Kim"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "University of Oregon"

authornote: |
     jcanfiel@uoregon.edu, Lundquist College of Business. 
     
     wkim4@uoregon.edu, Lundquist College of Business. 
     
     We would like to thank Professor Joseph Nese and participants at 2019 International Educational Associations Conference for helpful comments. 
  
abstract : |
  This paper reveals **gender** and **socioeconomic status** influence academic achievement, especially **math scores**. Along the way, it turns out that **teacher experience** is positively related to math scores such that the more teaching experience teachers have, the higher math scores students get. 
  
  
keywords          : "math scores, gender, socioeconomic status, teacher experience"

bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man, fleqn, noextraspace"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 8)

library(papaja)
library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)
library(rio)
library(here)

star <- import("~/Documents/workflow/data/star.csv")
```

# Introduction and Overview
Guo et al. [-@guo_et_al_2015] explore expectancy value in mathematics, gender and socioeconomic background as predictors of achievement. A multi-cohort study is used in the research [see @guo_et_al_2015]. Stoet and Geary [-@stoet_geary_2018] connects math with gender equality. Some other researchers study a similar topic [see @stoet_geary_2018]. Quinn and Cooc [-@quinn_cooc_2015] take these topics further by creating a model that predicts science achievement gaps by gender and race.  

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Data analysis

```{r summary table, results="asis"}
star %>% 
  group_by(sex, frl) %>% 
  summarize(math_mean = mean(tmathss),
            math_sd   = sd(tmathss),
            rdg_mean = mean(treadss),
            rdg_sd   = sd(treadss)) %>%
  kable(digits = 2,
            col.names = c("sex", "frl", "math_mean", "math_sd", "rdg_mean", "rdg_sd"), format = "latex") %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"), 
                font_size = 15)
```

As we can see above, girls perform better than boys in both math and reading. Regardless of gender, students who do not receive free lunch get higher average scores in both math and reading than students who receive free lunch.


```{r summary graph, results="asis"}
ggplot(star, aes(totexp, tmathss)) +
  geom_point(colour = "gray", size = 0.5) +
  geom_smooth(aes(color = frl), method = "lm") + 
  labs(title = "Relation between teacher experience and math scores", subtitle = "Separate regression lines displayed by free/reduced price lunch status", x = "Teacher Experience (years)", y = "Total Math Scale Score", color = "Lunch Status") +
  scale_color_discrete(labels=c("Paid Meals", "Free/Reduced Price Meals")) + theme_minimal() + 
  theme(legend.position = "bottom", legend.box = "horizontal") + 
  guides(col = guide_legend(reverse = TRUE))
```

Teacher experience has a weak positive relationship with math scores and it is true for both free/reduced meal students and paid meal students. 

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")

```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
